---
title: "Assignment 4"
author: "Kate"
date: "November 13, 2018"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r load packages and data, include=FALSE}

library(tidyverse)
library(pwr)
library(knitr)
library(kableExtra)
library(plotly)
library(ggrepel)
library(effsize)
library(pwr)
library(vcdExtra)
library(car)

size <- read_csv('lobster_size_abundance.csv')
trap <- read_csv('lobster_traps.csv')

```


```{r reformat data, include=FALSE}

# Reformat size data

size1 <- as.data.frame(size) %>% 
  filter(SIZE != -99999) %>% 
  expand.dft(., freq = 'COUNT') %>% 
  rename(
    year = YEAR,
    month = MONTH,
    date = DATE,
    site = SITE,
    lter_trans = SBC_LTER_TRANSECT,
    lob_trans = LOBSTER_TRANSECT,
    size = SIZE
  )

# Reformat trapping data

trap2 <- trap %>% 
  filter(SITE=='AQUE' |
           SITE=='CARP' |
           SITE=='IVEE' |
           SITE=='MOHK' |
           SITE=='NAPL') %>% 
  select(YEAR, SITE, TRAPS) %>% 
  rename(year = YEAR,
         site = SITE,
         traps = TRAPS)

```

###Part 1. Lobster abundance and fishing pressure (2012 - 2017)

**Ken:** Best way to depict these data?  1. Table of summary statistics for abundance and fishing data at each site 2. Figures of abundance at each site and fishing pressure (5 column graphs with abundance = column, and lines for fishing pressure)

```{r abundance: standardize for effort}

# Create data frame of effort: number of transects surveyed at each site for each year

effort <- size %>% 
  group_by(SITE, YEAR) %>% 
  summarize(effort = length(unique(SBC_LTER_TRANSECT)))

# Create data frame of abundance: total number of indv found at each site for each year

n <- size1 %>% 
  group_by(site, year) %>% 
  summarize(raw_n = length(size))

# Combine effort and abundance data frames

n2 <- full_join(n, effort, by = c('site'='SITE', 'year'='YEAR')) %>% 
  mutate(n = raw_n/effort)

# Add column for MPA/nonMPA
site <- unique(n2$site)
MPA <- c('N','N','Y','N','Y')
MPAs <- data.frame(site, MPA)

n3 <- full_join(n2, MPAs, by = 'site')

```


```{r abundance: scatter plot per site per year}

nsite_scatter <- ggplot(n3, aes(x = year, y = n)) +
  geom_point(aes(col = site)) +
  geom_line(aes(col = site)) +
  theme_minimal()

nsite_scatter

```


```{r abundance: scatter plot per year (all sites)}

n_year <- n %>% 
  group_by(year) %>% 
  summarize(total_n = sum(raw_n))

n_scatter <- ggplot(n_year, aes(x = year, y = total_n)) +
  geom_point() +
  geom_line() +
  theme_minimal()

n_scatter

```


```{r fishing: standardize for effort}

# Create data frame of effort: number of swaths surveyed at each site for each year

effort_trap <- trap2 %>% 
  group_by(site, year) %>% 
  summarize(effort = length(unique(traps)))

# Create data frame of abundance: total number of indv found at each site for each year

n <- size1 %>% 
  group_by(site, year) %>% 
  summarize(raw_n = length(size))

# Combine effort and abundance data frames

n2 <- full_join(n, effort, by = c('site'='SITE', 'year'='YEAR')) %>% 
  mutate(n = raw_n/effort)

# Add column for MPA/nonMPA
site <- unique(n2$site)
MPA <- c('N','N','Y','N','Y')
MPAs <- data.frame(site, MPA)

n3 <- full_join(n2, MPAs, by = 'site')

```


```{r fishing: scatter plot per site per year}

site_scatter <- ggplot(trap2, aes(x = year, y = traps)) +
  geom_point(aes(col = site)) +
  geom_line(aes(col = site)) +
  theme_minimal()

site_scatter

```

<br>

###Part 2. Mean lobster size by site in 2017


```{r}

size2017 <- size1 %>% 
  filter(year == "2017") %>%  # keeps only 2017
  select (year, site, size) # selects only 2017, the 5 sites, and size
  
```

```{r}
size_2017_hist<- ggplot(size2017, aes(x = size)) +
  geom_histogram(aes(fill = site)) +
  facet_wrap(~ site, scale = "free")

size_2017_hist
```

```{r}
qq_size2017<- ggplot(size2017, (aes(sample= size))) +
  geom_qq(aes(color = site)) +
  facet_wrap(~site, scale = "free")

qq_size2017

# based on qqplots and histograms observations for each group appear normally distrubuted. Also, n > 30 for all groups, so means will be normally distrubuted
```

```{r}
# If the largest sample variances is < 4x greater than the smallest sample variance, then these are "close enough" to consider variances equal

variances <- size2017 %>% 
  group_by(site) %>% 
  summarize(
    variance = var(size)
    )

variances # the largest variance (205.1)(IVEE) is definately not 4x greater than the smallest variance (86.1)(MOHK)

```
```{r}
# summary statistics for 2017 carapace size

size_sum <- size2017 %>% 
  group_by(site) %>% 
  summarize(
    Mean = round(mean(size), 2), # (mm)
    SD = round(sd(size), 2),
    n = length(size)) %>% 
  rename(Site = site) %>% 
  mutate(
    Site = case_when(
      Site == "AQUE" ~ "Arroyo Quemado", 
      Site == "CARP" ~ "Carpinteria", 
      Site == "IVEE" ~ "Isla Vista", 
      Site == "MOHK" ~ "Mohawk", 
      Site == "NAPL" ~ "Naples")) 
      
   
  
View(size_sum)
```

```{r}
table_caption <- '**Table 1.** **California spiny lobster (*Panulirus interruptus*) Carapace Length (mm) Recorded at five sites in the Santa Barbara Channel (2017).** Mean, standard deviation, and sample size of Carapace Length (mm) Recorded at five sites in the Santa Barbara Channel in 2017. Data retrieved from The Santa Barbara Coastal Long Term Ecological Research Project.' 

table_colnames <-c('Site', 'Mean', 'SD', 'n') 

size_anova_tab <-kable(size_sum,
                       caption = table_caption,
                       col.names = table_colnames,
                     align = c('l','c','c','c')) %>% 
  kable_styling(bootstrap_options = c("striped", "hover"),  
                full_width = FALSE)  


size_anova_tab
```

```{r}
# Levene Test 
# leveneTest() function in the 'car' package

levene_size2017 <- leveneTest(size ~ site, data = size2017)

levene_size2017

# H0: No difference in variances (variances are equal)
# HA: Variances are NoT equal

# There is a significant difference in variances across species
```

```{r}
# HO: The means across all groups are equal
# HA: At least two means differ

size2017_anova <- aov(size ~ site, data = size2017)

summary(size2017_anova)
```

```{r}
site_names <- c("Arroyo Quemado", "Carpinteria", "Isla Vista", "Mohawk", "Naples")

MPA_color <- c("white", "white", "gray60", "white", "gray60")

size_anova <- ggplot(size2017, aes(x = site, y = size)) +
  geom_boxplot(aes(fill = size)) +
  scale_fill_manual(values = MPA_color) +
  theme_classic() +
  scale_x_discrete(labels = site_names) +
  labs(x = "Site", y = "Carapace Length (mm) (n = 1668)") +
  theme(plot.title = element_text(hjust = 0.5))
  

size_anova

# color mpas
```



```{r}

# H0: There is no significant difference between means
# HA: There are 
size2017_ph <- TukeyHSD(size2017_anova)

size2017_ph
```

<br>

###Part 3. Changes in lobster size at MPA and non-MPA sites for 2012 and 2017

T test for MPAs
```{r}
size_2012_2017_MPA <- size1 %>% 
  filter(year == "2012" | year == "2017") %>%  # keeps only 2012 and 2017
  filter(site == "IVEE" | site == "NAPL") %>% #keeps the MPAs
  select (year, site, size) %>%  # selects '12 and '17, 5 sites, and size
  group_by(year)

View (size_2012_2017_MPA)
```

```{r, echo = FALSE}
# Individual dataframe for 2012

MPA2012 <- size_2012_2017_MPA  %>% 
  filter(year == "2012") %>%    # keeps only 2012
  pull (size)

View(MPA2012)
```

```{r, echo = FALSE}
# Individual dataframe for 2017

MPA2017 <- size_2012_2017_MPA  %>% 
  filter(year == "2017") %>%    # keeps only 2017
  pull (size)
  
View(MPA2017)
```

```{r, echo = FALSE}

# F-test for equal variances between 2012 and 2017

var.test(MPA2012,MPA2017, data = .) # F-test

# Variances are the same use a student's t

```

```{r, echo = FALSE}

# 2-sided, 2-sample t-test between 2012 and 2017

MPA_2012_2017_t <- t.test(MPA2012, MPA2017, data = ., var.equal = TRUE)

MPA_2012_2017_t
```

```{r, echo = FALSE}

# Effect size of the differences of size 2012 vs 2017

Cohen_d <- cohen.d(MPA2012, MPA2017)

Cohen_d

#(Cohen's d = 0.35)
# Absolute difference 2012 = 67.4 to 2017 = 72.2

```

T test for non-MPAs
```{r}
size_2012_2017_NON <- size1 %>% 
  filter(year == "2012" | year == "2017") %>%  # keeps only 2016
  filter(site == "MOHK" | site == "CARP" | site == "AQUE") %>% # keeps the non-MPAs
  select (year, site, size) %>%  # selects '12 and '17, 5 sites, and size
  arrange(site)

View (size_2012_2017_NON)
```

```{r, echo = FALSE}
# Individual dataframe for 2012

NON2012 <- size_2012_2017_NON  %>% 
  filter(year == "2012") %>%    # keeps only 2012
  pull (size) # keeps only size

View(NON2012)
```

```{r, echo = FALSE}
# Individual dataframe for 2017

NON2017 <- size_2012_2017_NON  %>% 
  filter(year == "2017") %>%    # keeps only 2017
  pull (size) # keeps only size
  
View(NON2017)
```

```{r, echo = FALSE}

# F-test for equal variances between 2012 and 2017

var.test(NON2012,NON2017, data = .) # F-test

# Variances are the same use a student's t

```

```{r, echo = FALSE}

# 2-sided, 2-sample t-test between 2012 and 2017

NON_2012_2017_t <- t.test(NON2012, NON2017, data = ., var.equal = TRUE)

NON_2012_2017_t
```

```{r, echo = FALSE}

# Effect size of the differences of size 2012 vs 2017

Cohen_d <- cohen.d(NON2012, NON2017)

Cohen_d

#(Cohen's d = 0.21)
# Absolute difference 2012 = 74.92 to 2017 = 72.3

```


###Part 4. Proportions of “legal” lobsters at the 5 sites in 2017

```{r}

size_n <- size1 %>% 
  filter(year == 2017) %>% 
  group_by(site) %>% 
  summarize(n = length(site))

size_legal <- size1 %>% 
  filter(year == 2017) %>% 
  mutate(legal = case_when(
    size>82.6 ~ 'Y',
    size<82.6 ~ 'N'
  )) %>% 
  group_by(site, legal) %>% 
  summarize(n_legal = length(legal)) %>% 
  filter(legal == 'Y') %>% 
  select(site, n_legal)

size_prop <- full_join(size_n, size_legal, by = 'site') %>% 
  mutate(prop_legal = n_legal/n)

```

The legal minimum carapace size for lobster is 82.6 mm. What proportion of observed lobsters at each site are above the legal minimum? Does that proportion differ significantly across the 5 sites? Note: We’ll be doing chi-square in labs next week, or go ahead with maximum resourcefulness and check out the chisq.test() function on your own!