---
title: "Assignment 4"
author: "Ken Gilliland, Kate Culhane, Sean Denny"
date: "November 13, 2018"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r load packages and data, include=FALSE}

library(tidyverse)
library(pwr)
library(knitr)
library(kableExtra)
library(plotly)
library(ggrepel)
library(effsize)
library(pwr)
library(vcdExtra)
library(car)
library(RColorBrewer)

size <- read_csv('lobster_size_abundance.csv')
trap <- read_csv('lobster_traps.csv')

```


```{r reformat data, include=FALSE}

# Create data frame for MPA information

site <- unique(size$SITE)
MPA <- c('Y','Y','N','N','N')
MPAs <- data.frame(site, MPA)

# Reformat size data so that each row is an individual lobster

size2 <- as.data.frame(size) %>% 
  filter(SIZE != -99999) %>% # filter out transects with no lobs
  expand.dft(., freq = 'COUNT') %>% # each row is an indv
  rename(
    year = YEAR,
    month = MONTH,
    date = DATE,
    site = SITE,
    lter_trans = SBC_LTER_TRANSECT,
    lob_trans = LOBSTER_TRANSECT,
    size = SIZE
  ) %>% 
  full_join(., MPAs, by = 'site')
size2$site <- factor(size2$site,
                     levels = c('AQUE', 'NAPL', 'IVEE', 'MOHK', 'CARP'))

# Simplify trapping data

trap2 <- full_join(trap, MPAs, by = c('SITE'='site')) %>% 
  filter(SITE=='AQUE' | # filter for our 5 sites of interest
           SITE=='CARP' |
           SITE=='IVEE' |
           SITE=='MOHK' |
           SITE=='NAPL') %>% 
  select(YEAR, SITE, MPA, TRAPS) %>% 
  rename(year = YEAR,
         site = SITE,
         traps = TRAPS)
trap2$site <- factor(trap2$site,
                     levels = c('AQUE', 'NAPL', 'IVEE', 'MOHK', 'CARP'))

```

###Part 1. Lobster abundance and fishing pressure (2012 - 2017)

**Ken:** Best way to depict these data?
1. Table of summary statistics for abundance and fishing data at each site
2. Figures of abundance at each site and fishing pressure (5 column graphs with abundance = column, and lines for fishing pressure)

```{r abundance: standardize for effort, include=FALSE}

# Create data frame of effort: number of transects surveyed at each site for each year

lob_effort <- size %>% 
  group_by(SITE, YEAR) %>% 
  summarize(effort = length(unique(SBC_LTER_TRANSECT)))

# Create data frame of abundance: total number of indv found at each site for each year

lob_n <- size2 %>% 
  group_by(site, year) %>% 
  summarize(raw_n = length(size))

# Combine effort and abundance data frames

abundance <- full_join(lob_n, lob_effort,
                by = c('site'='SITE', 'year'='YEAR')) %>% # combine effort and abundance data frames
  mutate(n = raw_n/effort) %>% # add column for abundance adjusted for effort
  full_join(., MPAs, by = 'site') # add MPA information

abundance$site <- factor(abundance$site,
                         levels = c('AQUE', 'NAPL', 'IVEE', 'MOHK', 'CARP'))

```


```{r FIG 2A - abundance: scatter plot per site over time, echo=FALSE}

# Scatter plot of lobster abundance over time at the 5 sites
# Abundance adjusted for survey effort

site_labs <- c('Arroyo Quemado', 'Naples*', 'Isla Vista*', 'Mohawk', 'Carpinteria')

n_site_scatter <- ggplot(abundance, aes(x = year, y = raw_n,
                                        group = site)) +
  geom_point(aes(col = site)) +
  geom_line(aes(col = site)) +
  scale_color_brewer(palette = 'Set2', labels = site_labs) +
  theme_classic() +
  labs(x = 'Year',
       y = 'Number of lobsters',
       title = 'Lobster abundance over time',
       col = 'Site')

n_site_scatter

```


```{r abundance: scatter plot over time, echo=FALSE}

# Scatter plot of lobster abundance over time
# All sites combined
# Not adjusted for survey effort (consistent over time)

n_scatter <- abundance %>% 
  group_by(year) %>% 
  summarize(total_n = sum(raw_n)) %>% 
  ggplot(aes(x = year, y = total_n)) +
    geom_point() +
    geom_line() +
    theme_minimal() +
    labs(x = 'Year',
         y = 'Total lobster abundance')

n_scatter

```


```{r fishing: standardize for effort, include=FALSE}

# Create data frame of effort: number of swaths surveyed at each site for each year

trap_effort <- trap2 %>% 
  group_by(site, year) %>% 
  summarize(effort = length(unique(traps)))

# Create data frame of abundance: total number of traps found at each site for each year

trap_n <- trap2 %>% 
  group_by(site, year) %>% 
  summarize(raw_traps = sum(traps))

# Combine effort and abundance data frames

fishing <- full_join(trap_n, trap_effort, 
                by = c('site', 'year')) %>% # combine effort and trap data frames
  mutate(traps = raw_traps/effort) %>% # add column for traps adjusted for effort
  full_join(., MPAs, by = 'site') # add MPA information

fishing$site <- factor(fishing$site,
                       levels = c('AQUE', 'NAPL', 'IVEE', 'MOHK', 'CARP'))

```


```{r FIG 2B - fishing: scatter plot per site over time, echo=FALSE}

# Create scatter plot of fishing pressure at each site over time
# Number of traps adjusting for survey effort

fishing2 <- fishing %>% 
  filter(site == 'AQUE' | site == 'MOHK' | site == 'CARP')

site_cols2 <- brewer.pal(5, 'Set2')[c(1,4,5)]
site_labs2 <- c('Arroyo Quemado', 'Mohawk', 'Carpinteria')

trap_site_scatter <- ggplot(fishing2, aes(x = year, y = raw_traps,
                                         group = site)) +
  geom_point(aes(col = site)) +
  geom_line(aes(col = site)) +
  scale_color_manual(values = site_cols2,
                     labels = site_labs2) +
  theme_classic() +
  labs(x = 'Year',
       y = 'Number of traps',
       title = 'Fishing pressure over time',
       col = 'Site') +
  theme(legend.position = 'none')

trap_site_scatter

```


```{r fishing: scatter plot over time, echo=FALSE}

# Scatter plot of fishing pressure over time
# All sites combined
# Not adjusted for survey effort

trap_scatter <- fishing %>% 
  group_by(year) %>% 
  summarize(total_traps = sum(raw_traps)) %>% 
  ggplot(aes(x = year, y = total_traps)) +
    geom_point() +
    geom_line() +
    theme_minimal() +
    labs(x = 'Year',
         y = 'Total number of traps')

trap_scatter

```


```{r compare abundance and fishing: scatter plot over time, echo=FALSE}

# Scatter plot comparing lobster abundance and trapping effort over time
# All sites combined
# Not adjusted for survey effort 

abundance2 <- abundance %>% 
  group_by(year) %>% 
  summarize(total_n = sum(raw_n))

fishing2 <- fishing %>% 
  group_by(year) %>% 
  summarize(total_traps = sum(raw_traps))

compare_scatter <- ggplot() +
  geom_point(dat=abundance2, aes(x = year, y = total_n)) +
  geom_line(dat=abundance2, aes(x = year, y = total_n)) +
  geom_point(dat=fishing2, aes(x = year, y = total_traps)) +
  geom_line(dat=fishing2, aes(x = year, y = total_traps)) +
  scale_y_continuous(limits = c(0,2500), sec.axis = dup_axis(name = 'Total number of traps')) +
  theme_minimal() +
  labs(x = 'Year',
       y = 'Total lobster abundance')

compare_scatter

```


```{r FIG 3 - compare abundance and fishing: scatter plot over time by MPA, echo=FALSE}

# Scatter plot comparing lobster abundance and trapping effort over time, for MPAs vs. non-MPAs
# Two sites combined for MPAs, three sites combined for non-MPAs
# NOT adjusted for survey effort 

abundance3 <- abundance %>% 
  group_by(year, MPA) %>% 
  summarize(total_n = sum(raw_n))

fishing3 <- fishing %>% 
  group_by(year, MPA) %>% 
  summarize(total_traps = sum(raw_traps))

MPA_labs <- c('non-MPA', 'MPA')
type_labs <- c('number of lobsters', 'number of traps')

compare_scatter2 <- ggplot() +
  geom_point(dat = abundance3, aes(x = year, y = total_n,
                                 group = MPA,
                                 col = MPA)) +
  geom_line(dat = abundance3, aes(x = year, y = total_n,
                                group = MPA,
                                col = MPA)) +
  geom_point(dat = fishing3, aes(x = year, y = total_traps/2,
                               group = MPA,
                               col = MPA)) +
  geom_line(dat = fishing3, aes(x = year, y = total_traps/2,
                              group = MPA,
                              col = MPA),
            linetype = 'dashed') +
  scale_y_continuous(sec.axis = sec_axis(~./2,
                                         name = 'Number of traps')) +
  scale_color_manual(values = c("grey70", "black"),
                     labels = MPA_labs) +
  theme_classic() +
  labs(x = 'Year',
       y = 'Number of lobsters',
       title = 'Lobster abundance and fishing pressure over time',
       col = '')

compare_scatter2

```

<br>

###Part 2. Mean lobster size by site in 2017

```{r size: data distribution and variance, include=FALSE}

# Data frame for 2017 data only
size2017 <- size2 %>% 
  filter(year == "2017") %>%  # keeps only 2017
  select (year, site, size, MPA)

# Histogram of size by site
size_2017_hist<- ggplot(size2017, aes(x = size)) +
  geom_histogram(aes(fill = site)) +
  facet_wrap(~ site, scale = "free")

size_2017_hist

# qq-plot of size by site
qq_size2017<- ggplot(size2017, (aes(sample= size))) +
  geom_qq(aes(color = site)) +
  facet_wrap(~site, scale = "free")

qq_size2017

# based on qqplots and histograms observations for each group appear normally distrubuted. Also, n > 30 for all groups, so means will be normally distrubuted



# Variance of size by site

# Levene Test 
# leveneTest() function in the 'car' package
# H0: No difference in variances (variances are equal)
# HA: Variances are NoT equal

levene_size2017 <- leveneTest(size ~ site, data = size2017)

levene_size2017
# There is a significant difference in variances across species

# Actual variances:
variances <- size2017 %>% 
  group_by(site) %>% 
  summarize(
    variance = var(size)
    )

variances

# the largest variance (205.1)(IVEE) is definately not 4x greater than the smallest variance (86.1)(MOHK)
# If the largest sample variances is < 4x greater than the smallest sample variance, then these are "close enough" to consider variances equal for ANOVA

```


```{r size: summary stats, echo=FALSE}

# summary statistics for 2017 carapace size

size_sum <- size2017 %>% 
  group_by(site) %>% 
  summarize(
    Mean = round(mean(size), 2), # (mm)
    SD = round(sd(size), 2),
    n = length(size)) %>% 
  rename(Site = site) %>% 
  mutate(
    Site = case_when(
      Site == "AQUE" ~ "Arroyo Quemado", 
      Site == "CARP" ~ "Carpinteria", 
      Site == "IVEE" ~ "Isla Vista", 
      Site == "MOHK" ~ "Mohawk", 
      Site == "NAPL" ~ "Naples")) 

table_caption <- '**Table 1.** **California spiny lobster (*Panulirus interruptus*) Carapace Length (mm) Recorded at five sites in the Santa Barbara Channel (2017).** Mean, standard deviation, and sample size of Carapace Length (mm) Recorded at five sites in the Santa Barbara Channel in 2017. Data retrieved from The Santa Barbara Coastal Long Term Ecological Research Project.' 

table_colnames <-c('Site', 'Mean', 'SD', 'n') 

size_anova_tab <-kable(size_sum,
                       caption = table_caption,
                       col.names = table_colnames,
                     align = c('l','c','c','c')) %>% 
  kable_styling(bootstrap_options = c("striped", "hover"),  
                full_width = FALSE)  


size_anova_tab

```


```{r size: ANOVA and post-hoc tests, include=FALSE}

# HO: The means across all groups are equal
# HA: At least two means differ

size2017_anova <- aov(size ~ site, data = size2017)

summary(size2017_anova)

# there's at least one sig diff!!

# post-hoc testing: Tukey's HSD 
size2017_ph <- TukeyHSD(size2017_anova)

size2017_ph
# significant differences between:
# 1. NAPL and CARP
# 2. NAPL and IVEE

# Summary table

size2017_summary <- size2017 %>% 
  group_by(site) %>% 
  summarize(mean = mean(size),
            sd = sd(size))

```

Satisfy the conditions of parametric tests (normal distribution, equal variance), so used ANOVA and Tukey's HSD. Only found significantdifferences between:

1. NAPL and CARP
2. NAPL and IVEE

```{r FIG 4 - size: boxplot, echo=FALSE}

site_names <- c("Arroyo Quemado", "Naples", "Isla Vista", "Mohawk", "Carpinteria")

size_anova <- ggplot(size2017, aes(x = site, y = size)) +
  geom_boxplot(aes(fill = MPA)) +
  scale_fill_manual(values = c('grey50', 'white')) +
  theme_classic() +
  scale_x_discrete(labels = site_names) +
  labs(x = "Site", y = "Carapace Length (mm) (n = 1668)") +
  theme(plot.title = element_text(hjust = 0.5))
  
size_anova

```

<br>

###Part 3. Changes in lobster size at MPA and non-MPA sites for 2012 and 2017

```{r MPA: data frames, include=FALSE}

# Create data frame for 2012, 2014, MPA data only
size_2012_2017_MPA <- size2 %>% 
  filter(year == "2012" | year == "2017") %>%  # keeps only 2012 and 2017
  filter(site == "IVEE" | site == "NAPL") %>% #keeps the MPAs
  select (year, site, size) %>%  # selects '12 and '17, 5 sites, and size
  group_by(year)

# Individual data frame for 2012

MPA2012 <- size_2012_2017_MPA  %>% 
  filter(year == "2012") %>%    # keeps only 2012
  pull (size)

# Individual dataframe for 2017

MPA2017 <- size_2012_2017_MPA  %>% 
  filter(year == "2017") %>%    # keeps only 2017
  pull (size)
  
```

```{r MPA: t-test , include = FALSE}

# F-test for equal variances between 2012 and 2017

var.test(MPA2012,MPA2017, data = .) # F-test

# Variances are the same --> use a student's t (override Welch approx)

# t-test
# 2-sided, 2-sample t-test between 2012 and 2017

MPA_2012_2017_t <- t.test(MPA2012, MPA2017, var.equal = TRUE)

MPA_2012_2017_t
# no significant difference

```


```{r non-MPA: data frames, include=FALSE}

# Create data frame for 2012, 2014, non-MPA data only

size_2012_2017_NON <- size2 %>% 
  filter(year == "2012" | year == "2017") %>%  # keeps only 2016
  filter(site == "MOHK" | site == "CARP" | site == "AQUE") %>% # keeps the non-MPAs
  select (year, site, size) %>%  # selects '12 and '17, 5 sites, and size
  arrange(site)

# Individual dataframe for 2012

NON2012 <- size_2012_2017_NON  %>% 
  filter(year == "2012") %>%    # keeps only 2012
  pull (size) # keeps only size

# Individual dataframe for 2017

NON2017 <- size_2012_2017_NON  %>% 
  filter(year == "2017") %>%    # keeps only 2017
  pull (size) # keeps only size
  
```

```{r non-MPA: t-test, include=FALSE}

# F-test for equal variances between 2012 and 2017

var.test(NON2012,NON2017, data = .) # F-test

# Variances are the same --> use a student's t (override Welch approx)

# t-test
# 2-sided, 2-sample t-test between 2012 and 2017

NON_2012_2017_t <- t.test(NON2012, NON2017, data = ., var.equal = TRUE)

NON_2012_2017_t
# there IS a sig diff

# Summary stats

size_2012_2017_NON_summary <- size_2012_2017_NON %>% 
  group_by(year) %>% 
  summarize(mean = mean(size),
            sd = sd(size))


```


```{r non-MPA: effect size, include=FALSE}

# Effect size of the differences of size 2012 vs 2017

Cohen_d <- cohen.d(NON2012, NON2017)

Cohen_d

#(Cohen's d = 0.21)
# Absolute difference 2012 = 74.92 to 2017 = 72.3

```

Variances were equal, so used student's t-test to compare 2012 and 2017 and Cohen's d to calculate effect size. Results:

**MPAs:** NO significant change in sizes from 2012 to 2017.

**non-MPAs:** significant change change in sizes from 2012 to 2017

- effect size is small
- actual difference: 2012 = 74.92, 2017 = 72.3


###Part 4. Proportions of “legal” lobsters at the 5 sites in 2017

```{r legality: data frame, echo=FALSE}

# Make contingency table of counts of legal and illegal lobs per site
size_legal <- size2 %>% 
  filter(year == 2017) %>% 
  mutate(legal = case_when( # add column for legality (Y/N)
    size>82.6 ~ 'Y',
    size<82.6 ~ 'N'
  )) %>% 
  select(site, legal)

legal <- table(size_legal) # turn into a contigency table

# Make a data frame with counts and proportions
legal_df <- as_data_frame(legal) # data frame from the contingency table
legalN <- legal_df[1:5, c(1,3)] # first 5 rows are illegal lobs
legalY <- legal_df[6:10, c(1,3)] # last 5 rows are legal lobs
legal2 <- full_join(legalN, legalY, by = 'site') %>% # combine 
  rename(Y = n.y,
         N = n.x) %>% 
  mutate(n = Y+N, # add column for total number of lobs
         propN = N/n, # prop of illegal lobs
         propY = Y/n) # prop of legal lobs

legal2

```


```{r legality: chi-square, include=FALSE}

# Chi-square test

size_chi <- chisq.test(size_legal$site, size_legal$legal)

size_chi
# significant association between site and size

```

Ran chi-square on site vs. size and found a significant difference.